Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-01-27 text/src/article.tex
Revise last section of implementation

The choice of nucleotide sequences comprising the genome scaffold is specified
in YAML \cite{yaml}. YAML is a language format for data using whitespace and
identation to produce a machine readable structure. As a standardised data
format third-party developers have the option to generate a genome scaffold
using any programming language for which a YAML parser exists. Furthermore YAML
formatted scaffold files can be validated using third party tools such as
Kwalify\cite{kwalify}.\pb

The initial assembly of sequencing reads may result in an incomplete genome
build. Comparison of contigs to reference genomes and the generation of
additional sequence from PCR translates to genome finishing as an ongoing
process. Therefore the scaffold file should be simple to edit manually in
additon to being computationally tractable. This requirement for the scaffold
files to be modifable in a simple simple text editor is the second reason YAML
was selected as the format for scaffold files.\pb

\subsection*{Parsing the YAML scaffold} %{{{2

The entries in the YAML scaffold are defined as a list. Each entry in the list
corresponds to a region sequence in the final output sequence which follows the
previous entry in the scaffold and precedes the following entry.  Each
attribute defined for each entry in the list allows specifications for the
entry such as if any starting sequence should be clipped from the beginning or
end, or if the sequence should be reversed.\pb

INSERT IMAGE OF SCAFFOLD GRAPH AND RESULTING OUTPUT.

Each stretch of sequence in the genome scaffold is defined as a list entry in
the YAML scaffold file. When parsing the scaffold file Scaffolder iterates over
each scaffold entry to create an instance of the Scaffolder::Region class.\pb

The Region class is inherited by all types of sequence in the scaffold. The
Region provides helper methods common to the tasks in manipulating scaffold
sequence, subsequencing and reversing. In particular the region class has
a class annotation method\cite{class-annotations} which allows attributes in
the scaffold method to be defined in the class. This negates having to write
a custom method to deal with each scaffold variable.  Furthermore this makes
the process of transferring attributes from the scaffold file to each region
instance much simpler - each yaml attribute corresponds to a method call on the
instantiated class. The scaffold file effectively becomes a DSL for populating
the scaffold region instances while still maintaining the machine readable data
qualities of YAML.\pb

The region class provides a sequence hook method. This method returns by
default the sequence corresponding to the fasta sequence defined by the source
tag which is then sublclassed or reversed if necessary. Subclasses of the
region class should override this method to provide custom functionality in
manipulating the source sequence. For example, the Scaffolder::Region::Sequence
class overrides the sequence method add any sequence inserts specified in the
scaffold file.\pb

>>>>

The choice of nucleotide sequences comprising the genome scaffold is specified
in YAML \cite{yaml}. YAML is a language format for data using white-space and
indentation to produce a machine readable structure. As a standardised data
format third-party developers have the option to generate a genome scaffold
using any programming language for which a YAML parser exists. Furthermore YAML
formatted scaffold files can be validated using third party tools such as
Kwalify\cite{kwalify}.\pb

The initial assembly of sequencing reads may result in an incomplete genome
build. Comparison of contigs to reference genomes and the generation of
additional sequence from PCR translates to genome finishing as an ongoing
process. Therefore the scaffold file should be simple to edit manually in
addition to being computationally tractable. This requirement for the scaffold
files to be modifiable in a simple simple text editor is the second reason YAML
was selected as the format for scaffold files.\pb

\subsection*{Processing the scaffold file} %{{{2

The form of the scaffold file is a list of entries. Each entry corresponds to
a length of sequence in the the draft genome sequence. When generating the
genome sequence from the scaffold file each entry is processed in order. Each
of the nucleotide sequence specified by the scaffold file entries are
concatenated together to produce the draft genome sequence. Each entry in the
scaffold file may have attributes which define whether a sub-sequence or the
reverse complement of the sequence should be used is the scaffold. The types of
attributes are outlined in the Results and Discussion section below.\pb

As the scaffold file is processed a corresponding ruby object instance is
created to correspond to an entry in the scaffold file. as an instance of
a subclass of Scaffolder::Region.  Scaffolder::Region is the super class for
all region entries in the scaffold.  This Region class provides the helper
methods for sub-sequencing and reversing complementing a sequence.  The Region
class also provides a class annotation method\cite{class-annotations} which
simplifies defining the attributes for each scaffold entry. Possible attributes
for scaffold entries can be added by using this helper method. \pb

An noteworthy method of the Scaffolder::Region class is the ``raw_sequence''
method. In the Scaffolder::Region class method returns by the sequence defined
by the ``source'' tag in the scaffold file. This sequence is then sub-sequenced
or reversed complemented if necessary. Subclasses of the Scaffolder::Region
class override this method to additional manipulate the nucleotide sequence.
For example, the Scaffolder::Region::Sequence class overrides the sequence
method to insert any specified inserts specified in the scaffold file before
this sequence is returned up the method chain.\pb
